
```{r}
library(readxl)
ridership <- read_excel("C:/Users/luenx/OneDrive/Desktop/UMP/UMP/Sem 6/group project/final ridership.xlsx")
ridership
```

```{r}
x1 <- 'Failure RS'
x2 <- 'Failure TNM'
x3 <- 'Failure POW'
x4 <- 'Failure SIG'
x5 <- 'Relevant Failure'
y1 <- 'Service Downtime'
y2 <- 'Total Ridership'
df <- ridership[c(x1, x2, x3, x4, x5, y1, y2)]
df
```
```{r}
# Replace column names
names(df) <- c("x1", "x2", "x3","x4","x5","y1","y2")
df

```



```{r}
# Check for null values in the entire data frame
is.na(df)
```

```{r}
df<- na.omit(df)
df
```

```{r}
#Descriptive Statistics
sapply(df,mean)  #mean
sapply(df,sd)    #standard deviation
sapply(df,var)   #variance
A<-cov(df)   #variance-covariance matrix
A
summary(df) #other statistics
```
```{r}
#Correlation matrix 
 cormatrix<-cor(df) 
 round(cormatrix,4)

#Multiple Scatter Plot matrix
library(car)
scatterplotMatrix(df)
```
X1 vs Others: X1 shows a positive linear relationship with X3, X4, and X5, and a nonlinear trend with y1 and y2.
X2 vs Others: X2 shows a non-linear relationship with other variables. It seems to have a distinctive behavior compared to other variables, possibly indicating some form of threshold effect or non-linear interaction.
X3 vs Others: X3 shows a positive linear relationship with X1, X4, and X5. There is a slightly curved relationship with y1 and y2.
X4 vs Others: X4 shows a strong positive linear relationship with X1, X3, and X5. The relationship with y1 and y2 is also positive but not as strong.
X5 vs Others: X5 has a positive linear relationship with X1, X3, and X4. The relationships with y1 and y2 are less clear and might be non-linear.
y1 and y2: The relationship between y1 and y2 is complex, possibly indicating an interaction or a more intricate dependency structure.





```{r}
#Multivariate Regression Model
 y<-as.matrix(df[,6:7])
 x<-as.matrix(df[,1:5])
 mvlr<-lm(y~x)
 summary(mvlr)
```
For Response y1,
^y1 = -7.5563 + 0.2717x1 + 2.9839x2 + 0.3473x3 - 0.7783x4 + 17.2527x5

x1 and x5 are significant variables at alpha=0.05.

Based on the P-value=0.000 < 0.05, this model is significant where at least one of the independent variables related to dependent variables. 


For Response y2,
^y2 = 4307376 - 7671x1 - 65936x2 + 95022x3 + 16290x4 + 157459x5

None of the variable is significant variables at alpha=0.05 except for intercept.

Based on the P-value=0.2762 > 0.05, this model is not significant where none of the independent variables is related to the dependent variables. 







```{r}
#Multivariate Test Statistics
 library(car)
 mv1r1<-lm(y~x1+x2+x3+x4+x5,df)
 Anova(mv1r1)
```
-Based on the result, x5 has the p-value <0.05.
-Therefore, x5 is the significant predictor to this multivariate model.


```{r}
#Multivariate Test Statistics
 library(car)
 mv1r1<-lm(y~x1+x2+x3+x4+x5,df)
 Anova(mv1r1)
 mv1r2<-lm(y~x5,df)
 anova(mv1r1,mv1r2)
```
-The p-value = 0.00003 < 0.05, hence it is significant.

-Hence, there is enough evidence that the model with one predictor fits as well as the model with five predictors at 5% significance level.



#Option 2 
```{r}
#Multivariate Test Statistics 
 library(car)
 mv1r1<-lm(y~x1+x2+x3+x4+x5,df)
 Anova(mv1r1)
 linearHypothesis(mv1r1,hypothesis.matrix=c("x1=0"))
```
-P-value of all the tests such as Pillai, Wilks, Hotelling-Lawley and Roy are not significant at 5% significant level.
-Hence, the P-value does not provide enough evidence that the model with four predictors fits as well as the model with five predictors at 5% significance level. Thus, the model with four predictors cannot be used for prediction.

^y1 = -7.556e+00 + 2.717e-01x1 - 2.984e+00x2 + 3.473e-01x3 + -7.783e-01x4 + 1.725e+01x5 
^y2 = 4,307e+06 + 7.671e+03x1 - 6.594e+04x2 + 9.502e+04x3 + 1.629e+04x4 + 1.575e+05x5 




```{r}
lm(y~x1+x2+x3+x4+x5,df)
```
Interpretation:
^y1 = -7.556e+00 + 2.717e-01x1 - 2.984e+00x2 + 3.473e-01x3 + -7.783e-01x4 + 1.725e+01x5 

-This model can be used for prediction. 



```{r}
####model checking###
#install.packages("car")

# Load the car package
library(car)

yhat<-predict(mvlr) #predicted y
epsilon<-residuals(mvlr)   #residuals
cbind(yhat,epsilon)
```


```{r}
scatterplotMatrix(epsilon) #Linearity
```


```{r}
df<-data.frame(epsilon)   #change into data frame to construct QQ-plot (normality)
qqPlot(df$y1)
qqPlot(df$y2)
```
-Outliers exist for y1 which are 2 and 12. 
-Outliers exist for y2 which are 7 and 16.


```{r}
#HOMOSCEDASTICITY / CONSTANT VARIANCE
plot(df$y1)
plot(df$y2)

B<-cov(df)
B 
```
-Homoscedasticity does not occur as the residuals are not equally distributed across the regression line.


#EXTRA CODING
```{r}
#library(lmtest)
# Fit the linear model
#mv1r1 <- lm(y ~ x1 + x2 + x3 + x4 + x5, data = df)

# Perform the Breusch-Pagan test
#bptest(mv1r1)

```
If the p-value is small (typically less than 0.05), it suggests the presence of heteroscedasticity.
